---
title: "A Short Introduction to profoc"
author: Jonathan Berrisch
date: "`r Sys.Date()`"
bibliography:
  - ../inst/bib/profoc.bib
output:
  rmarkdown::html_vignette:
    number_sections: yes
    toc: no
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(profoc)
```

# Introduction

The R package **profoc** offers a user-friendly way to apply online-learning algorithms for forecast combination. It implements multiple algorithms as well as loss functions. The algorithms are implemented in C++ and can be used in R via Rcpp [@eddelbuettel2013springer]. The main function `online()` offers a high-level interface to the learning algorithms and their extensions. It is a wrapper to the C++ class `conline` which implements the online-learning algorithms. The class functions as a low-level interface. We exposed it for advanced users who need to adjust specific details. 

In this introductory vignette, we demonstrate the use of `online()` to run the core algorithms. We also show how the results can be inspected. Various popular extensions of the core algorithms, methods for deployment in production, and the low-level interface will be covered in separate vignettes.

## Online Learning